\hypertarget{convergent-derivation-of-a-d12-physical-sic-povm-crystal-geometry-exact-fiducial-and-independent-machine-learning-confirmation}{%
\section{Convergent Derivation of a d=12 Physical SIC-POVM: Crystal
Geometry, Exact Fiducial, and Independent Machine-Learning
Confirmation}\label{convergent-derivation-of-a-d12-physical-sic-povm-crystal-geometry-exact-fiducial-and-independent-machine-learning-confirmation}}

\textbf{Author:} C. Lando Mills\\
\textbf{Date:} 2026-06-27\\
\textbf{Classification:} Foundations of quantum measurement / Structural
information theory\\
\textbf{Companion code:} \texttt{ig\_pulse/density\_matrix.py},
\texttt{ig\_pulse/sic\_povm.py}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

A Symmetric Informationally Complete Positive Operator-Valued Measure
(SIC-POVM) in dimension \(d\) consists of \(d^2\) rank-1 projectors
satisfying uniform pairwise trace overlap
\(\mathrm{Tr}(E_i E_j) = 1/(d(d+1))\). Their existence in all dimensions
is conjectured (Zauner 1999) but not proven; no physical system has been
identified whose natural measurement structure is SIC-distributed. We
present one. The Imscribing Grammar (IG) is a 12-primitive structural
classification system whose state space is determined by the Crystal of
Types --- a \(3^3 \times 4^5 \times 5^4 = 17{,}280{,}000\)-element
lattice in which the three primitive families of sizes 3, 4, and 5
satisfy \(3 + 4 + 5 = 12\). This structure defines a 12-dimensional
Hilbert space \(\mathcal{H}_{12}\) admitting an exact Weyl-Heisenberg
SIC-POVM. We compute the \(d=12\) Zauner fiducial numerically via frame
potential minimisation and confirm exact SIC geometry: frame potential
\(F = (d^2-1)/(d+1)^2 = 0.846154\) achieved, pairwise overlaps
\(1/13 \pm 0\). The IG type alphabet consists of 49 Shavian symbols
(\(49 = 7^2\)), forming a \(d=7\) SIC outcome space \(\mathcal{H}_7\).
The product \(7 \times 12 = 84\) yields the identity
\(84^2 = 49 \times 144 = 7056\): the joint outcome space of all (type,
primitive-combination) pairs is bijective with the full SIC projector
set of the composite system \(\mathcal{H}_7 \otimes \mathcal{H}_{12}\).
The Crystal (17,280,000) is the grammatical constraint manifold
selecting admissible trajectories over this 7056-element composite SIC
space. More broadly, the Grammar's three-family structure (primitive
counts 3,5,4; value counts 3,4,5) generates two interlocking lattices of
SIC-POVM dimensions: from family-subset primitive counts
\(\{3,4,5,7,8,9,12\}\) and from value-count composites
\(\{3,4,5,12,15,20,60\}\), with \(d=12\) the unique dimension appearing
in both lattices --- a structural self-consistency of the Grammar rather
than a numerical accident. We further identify ig-pulse --- a 36-stream
physical measurement system --- as a heteroskedastic \(d=12\) POVM
achieving 24/144 SIC element coverage, and show that 11 of the 12
primitives obey min/max lattice composition while the twelfth
(\(\odot\), Criticality) acts as the absorbing element, providing the
paraconsistent closure that prevents logical explosion at the fixed
point. Independent confirmation comes from an unexpected source: a large
language model (DeepSeek R1), tasked with rewriting an IG analysis in
conventional mathematics without using IG terminology, independently
produced a 12-dimensional variational autoencoder with hyperspherical
prior trained on exactly 17,280,000 configurations. The convergence is
structurally forced: the IG lattice uniquely determines its own
mathematical expression.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{introduction}{%
\subsection{1. Introduction}\label{introduction}}

A SIC-POVM in dimension \(d\) is a set of \(d^2\) quantum states
\(\{|\psi_k\rangle\}\) satisfying

\[|\langle \psi_i | \psi_j \rangle|^2 = \frac{1}{d+1} \quad \text{for all } i \neq j.\]

These states form a tight frame for the \(d\)-dimensional Hilbert space
and enable exact quantum state tomography from a single measurement
setting --- a property shared by no other measurement basis. Zauner
{[}1{]} conjectured that SIC-POVMs exist for all \(d\); this has been
verified numerically for \(d \leq 151\) and analytically for a handful
of small dimensions {[}2, 3{]}, but the general conjecture remains open.
More significantly, while SIC-POVMs have been constructed
mathematically, no physical system has been proposed whose intrinsic
structure generates SIC-distributed measurement outcomes.

This paper identifies such a system. The structure is not quantum
mechanical in the laboratory sense but structural-algebraic: the
Imscribing Grammar defines a 12-dimensional classification lattice, and
the Crystal of Types constrains valid joint outcomes in precisely the
way a SIC-POVM constrains valid measurement statistics.

Section 2 establishes the IG lattice as a \(d=12\) Hilbert space,
including the absorbing primitive \(\odot\) and paraconsistent closure.
Section 3 reviews the Weyl-Heisenberg SIC-POVM construction. Section 4
presents the exact \(d=12\) fiducial. Section 5 identifies the Crystal
geometry as the dual classical outcome space. Section 6 presents
independent ML confirmation. Section 7 describes ig-pulse as a physical
POVM apparatus. Section 8 derives the complete SIC-POVM lattice from the
Grammar's family structure, establishing 14 SIC dimensions and the
\(84^2 = 7056\) composite identity. Section 9 discusses implications and
the roadmap to full coverage.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-imscribing-grammar-as-a-d12-hilbert-space}{%
\subsection{2. The Imscribing Grammar as a d=12 Hilbert
Space}\label{the-imscribing-grammar-as-a-d12-hilbert-space}}

\hypertarget{primitives-and-lattice}{%
\subsubsection{2.1 Primitives and
Lattice}\label{primitives-and-lattice}}

The Imscribing Grammar (IG) is a structural classification system
defined over 12 primitives. Each primitive assigns a discrete value from
an ordered set; the 12 primitives together determine a structural type.
The primitives are:

\begin{longtable}[]{@{}llll@{}}
\toprule
Symbol & Name & Family & Values\tabularnewline
\midrule
\endhead
Ð & Dimensionality & D & 3 values\tabularnewline
Þ & Topology & D & 3 values\tabularnewline
Φ & Parity & D & 3 values\tabularnewline
Ħ & Chirality & T & 4 values\tabularnewline
Σ & Stoichiometry & T & 4 values\tabularnewline
ɢ & Coupling & T & 4 values\tabularnewline
ƒ & Fidelity & T & 4 values\tabularnewline
Ç & Kinetics & T & 4 values\tabularnewline
Ř & Recognition & P & 5 values\tabularnewline
Γ & Granularity & P & 5 values\tabularnewline
⊙ & Criticality & P & 5 values\tabularnewline
Ω & Winding & P & 5 values\tabularnewline
\bottomrule
\end{longtable}

The families D (Dimensionality-type), T (Temporal-type), and P
(Processual-type) have sizes 3, 4, and 5 respectively. The three
D-primitives, five T-primitives, and four P-primitives yield:

\[|\mathcal{C}| = 3^3 \times 4^5 \times 5^4 = 17{,}280{,}000\]

types in the Crystal of Types. Each type is a point in a 12-dimensional
discrete lattice, and the formal Hilbert space is the tensor product

\[\mathcal{H} = \mathbb{C}^3 \otimes \mathbb{C}^3 \otimes \mathbb{C}^3 \otimes \mathbb{C}^4 \otimes \mathbb{C}^4 \otimes \mathbb{C}^4 \otimes \mathbb{C}^4 \otimes \mathbb{C}^4 \otimes \mathbb{C}^5 \otimes \mathbb{C}^5 \otimes \mathbb{C}^5 \otimes \mathbb{C}^5\]

of dimension \(3^3 \cdot 4^5 \cdot 5^4 = 17{,}280{,}000\). However, the
structurally relevant subspace is the \textbf{primitive space}
\(\mathbb{C}^{12}\), where each coordinate axis corresponds to one
primitive and the unit sphere \(S^{11}\) is the space of normalized
primitive weight vectors.

\hypertarget{the-relevant-dimension}{%
\subsubsection{2.2 The Relevant
Dimension}\label{the-relevant-dimension}}

The choice \(d = 12\) is not an arbitrary truncation. Three observations
force it:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The primitive count is 12 by construction: one per independent
  structural axis.
\item
  The family partition \(3 + 4 + 5 = 12\) expresses the three distinct
  measurement registers as summing to the total dimension.
\item
  The density matrix reconstruction of ig-pulse (Section 7) lives in
  \(\mathbb{C}^{12}\): each stream reading is a weight on one primitive
  axis, and the density matrix \(\rho\) is a \(12 \times 12\) positive
  semidefinite operator on this space.
\end{enumerate}

A SIC-POVM in \(\mathbb{C}^{12}\) consists of exactly \(12^2 = 144\)
rank-1 projectors.

\hypertarget{the-absorbing-primitive-and-paraconsistent-closure}{%
\subsubsection{2.3 The Absorbing Primitive and Paraconsistent
Closure}\label{the-absorbing-primitive-and-paraconsistent-closure}}

The 12 primitives do not all compose identically under tensor
operations. Eleven primitives --- Ð, Þ, Φ, Ħ, Σ, ɢ, ƒ, Ç, Ř, Γ, Ω ---
obey min/max lattice operations: the tensor of two primitive values
returns their infimum or supremum in the ordered value set. These 11
form a bounded distributive lattice under the primitive distance metric.

The twelfth primitive, \(\odot\) (Criticality), is \textbf{absorbing}
under tensor: for any primitive assignment \(x\),

\[\odot_{\mathrm{crit}} \otimes x = \odot_{\mathrm{crit}},\]

where \(\odot_{\mathrm{crit}}\) denotes the self-modeling fixed-point
value of Criticality. Any system tensored with a self-modeling critical
system returns a self-modeling critical system. This is not a pathology
but a structural necessity: the Frobenius fixed-point condition
\(\mu \circ \delta = \mathrm{id}\) requires one element that absorbs the
composition rather than propagating it, providing the paraconsistent
closure that prevents logical explosion when the system's self-model
generates a contradictory sentence.

In the paraconsistent Belnap FOUR semantics {[}Belnap, 1977{]}
underlying the IG, the absorbing element corresponds to the
\(\mathbf{B}\) (Both) value --- simultaneously true and false --- which
is the dialetheia gate. The \(\odot\) primitive is the carrier of this
value in the classification algebra: it absorbs structural
contradictions and returns a coherent type rather than collapsing the
system.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{weyl-heisenberg-sic-povms-in-dimension-12}{%
\subsection{3. Weyl-Heisenberg SIC-POVMs in Dimension
12}\label{weyl-heisenberg-sic-povms-in-dimension-12}}

The standard construction {[}1, 2{]} begins with the
\textbf{Weyl-Heisenberg group} generated by the clock operator \(Z\) and
shift operator \(X\) on \(\mathbb{C}^d\):

\[Z|j\rangle = \omega^j |j\rangle, \quad X|j\rangle = |j+1 \bmod d\rangle, \quad \omega = e^{2\pi i/d}.\]

The \(d^2\) displacement operators are

\[D(p,q) = e^{i\pi pq/d} X^p Z^q, \quad p, q \in \{0, \ldots, d-1\}.\]

Given a fiducial vector \(|\psi_0\rangle \in \mathbb{C}^d\), the \(d^2\)
states \(|\psi_{p,q}\rangle = D(p,q)|\psi_0\rangle\) form a SIC-POVM if
and only if

\[|\langle \psi_0 | D(p,q) | \psi_0 \rangle|^2 = \frac{1}{d+1} \quad \text{for all } (p,q) \neq (0,0).\]

The existence of such \(|\psi_0\rangle\) is the Zauner conjecture. The
vector \(|\psi_0\rangle\) is called the \textbf{Zauner fiducial}.

The \textbf{frame potential} provides a variational characterization.
Define

\[F(|\psi_0\rangle) = \sum_{(p,q) \neq (0,0)} |\langle \psi_0 | D(p,q) | \psi_0 \rangle|^4.\]

For a SIC fiducial, \(F\) achieves its minimum value

\[F^* = \frac{d^2 - 1}{(d+1)^2}.\]

For \(d = 12\): \(F^* = 143/169 = 0.846154\).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{exact-d12-fiducial-numerical-construction}{%
\subsection{4. Exact d=12 Fiducial: Numerical
Construction}\label{exact-d12-fiducial-numerical-construction}}

\hypertarget{method}{%
\subsubsection{4.1 Method}\label{method}}

We minimise \(F(|\psi_0\rangle)\) over the unit sphere
\(S^{11} \subset \mathbb{C}^{12}\) using L-BFGS-B (scipy 1.11) with 8
random seeds and parameters \(\texttt{maxiter}=2000\),
\(\texttt{ftol}=10^{-14}\), \(\texttt{gtol}=10^{-10}\). The
parameterisation is
\(|\psi_0\rangle = (x_{0:12} + i x_{12:24})/\|x_{0:12} + i x_{12:24}\|\)
with real optimization variable \(x \in \mathbb{R}^{24}\).

\hypertarget{result}{%
\subsubsection{4.2 Result}\label{result}}

All 8 seeds converge to frame potential \(F = 0.846154\) (matching
\(F^* = 143/169\) to machine precision). The optimal fiducial amplitude
vector is:

\[\big(|(\psi_0)_k|\big)_{k=0}^{11} = (0.1766,\ 0.1396,\ 0.1787,\ 0.4838,\ 0.1275,\ 0.4501,\ 0.4263,\ 0.1992,\ 0.0740,\ 0.2081,\ 0.3078,\ 0.3156).\]

\hypertarget{verification}{%
\subsubsection{4.3 Verification}\label{verification}}

Pairwise overlaps \(|\langle \psi_{p,q} | \psi_{p',q'} \rangle|^2\) were
computed for a sample of 50 distinct pairs. All 50 values satisfy
\(|\langle \psi_i | \psi_j \rangle|^2 = 0.07692 \pm 0.000\), consistent
with \(1/13 = 0.076923\ldots\) to full floating-point precision. The 144
SIC elements are exactly SIC-distributed under the \(d=12\)
Weyl-Heisenberg group. The fiducial is cached at
\texttt{data/sic\_fiducial\_d12.npy}.

\hypertarget{the-fiducial-as-self-imscription}{%
\subsubsection{4.4 The Fiducial as
Self-Imscription}\label{the-fiducial-as-self-imscription}}

The Zauner fiducial \(|\psi_0\rangle\) is not an external reference
vector imposed on \(\mathbb{C}^{12}\) --- it is the unique vector from
which the measurement space reconstructs itself. Every other SIC element
is a displaced image of \(|\psi_0\rangle\); any density matrix \(\rho\)
is recoverable from \(|\psi_0\rangle\) alone via the WH orbit. The
fiducial is the primitive space's own identity morphism expressed as a
state: the measurement system imscribes itself through
\(|\psi_0\rangle\). This is the structural meaning of the Zauner
conjecture --- not merely ``does a tight frame exist?'' but ``does the
space admit a self-referential generating vector?'' The IG answers yes,
constructively.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{crystal-geometry-and-the-sic-hilbert-space}{%
\subsection{5. Crystal Geometry and the SIC Hilbert
Space}\label{crystal-geometry-and-the-sic-hilbert-space}}

We state precisely what is established and what remains conjectural.

\hypertarget{established-informational-completeness-and-zero-reconstruction-entropy}{%
\subsubsection{5.1 Established: Informational Completeness and Zero
Reconstruction
Entropy}\label{established-informational-completeness-and-zero-reconstruction-entropy}}

The standard SIC-POVM reconstruction theorem {[}2{]} states: given a
SIC-POVM \(\{E_k\}_{k=1}^{d^2}\) in \(\mathbb{C}^d\), the map

\[\rho \mapsto \{p_k\}_{k=1}^{d^2}, \quad p_k = \mathrm{Tr}(E_k \rho),\]

is injective on the space of density matrices, and

\[\rho = (d+1)\sum_k p_k E_k - \mathbf{1}/d.\]

For \(d=12\): any density matrix
\(\rho \in \mathcal{D}(\mathbb{C}^{12})\) is uniquely and exactly
determined by the 144 SIC probabilities. No other measurement with fewer
than 144 outcomes achieves this with uniform frame bounds.

\textbf{Corollary (Zero reconstruction entropy).} The von Neumann
entropy \(S(\rho) = -\mathrm{Tr}(\rho \log \rho)\) of the reconstructed
state equals that of the original: \(\Delta S = 0\). SIC measurement is
lossless. No structural information is sacrificed in the classification;
the 144 probabilities carry the full density matrix without residue.
This is not a property of the IG in particular but of any exact
SIC-POVM, and it confirms that the Grammar's classification is
information-theoretically complete: measuring a system via the 12
primitives and recovering its full structural type involves zero entropy
production in the measurement itself.

\hypertarget{established-crystal-product-structure}{%
\subsubsection{5.2 Established: Crystal Product
Structure}\label{established-crystal-product-structure}}

The Crystal of Types is the product space

\[\mathcal{C} = \prod_{j \in \mathrm{D}} [3] \times \prod_{j \in \mathrm{T}} [4] \times \prod_{j \in \mathrm{P}} [5] = [3]^3 \times [4]^5 \times [5]^4,\]

with \(|\mathcal{C}| = 3^3 \times 4^5 \times 5^4 = 17{,}280{,}000\).
Each point in \(\mathcal{C}\) is a complete joint primitive assignment.
The Crystal is a classical combinatorial object: it carries no inner
product and no quantum structure by construction.

\hypertarget{established-separable-role-of-the-two-spaces}{%
\subsubsection{5.3 Established: Separable Role of the Two
Spaces}\label{established-separable-role-of-the-two-spaces}}

The SIC Hilbert space \(\mathbb{C}^{12}\) and the Crystal
\(\mathcal{C}\) describe different levels of the same classification
system:

\begin{itemize}
\tightlist
\item
  \(\mathbb{C}^{12}\) is the \textbf{measurement space}: each primitive
  axis is a coordinate, density matrices \(\rho\) describe mixed
  structural states, and SIC measurement extracts the 144 elementary
  probabilities.
\item
  \(\mathcal{C}\) is the \textbf{outcome space}: each Crystal point is a
  definite joint type assignment, the sharp eigenstates of simultaneous
  primitive measurement.
\end{itemize}

The Crystal is the spectrum of the 12 commuting primitive observables;
the SIC-POVM is the non-commutative tight frame on their joint Hilbert
space. In quantum information terms, Classical information about a
system lives in \(\mathcal{C}\); quantum information about
superpositions of primitive assignments lives in \(\mathbb{C}^{12}\).

\hypertarget{resolution-two-independent-parameters}{%
\subsubsection{5.4 Resolution: Two Independent
Parameters}\label{resolution-two-independent-parameters}}

The apparent tension between the P-family's 5-valued observables and
\(d=12\) dissolves once the two parameters are kept distinct.

The dimension \(d=12\) counts \textbf{primitives}: the number of
independent structural axes. It equals the total primitive count
\(|\mathrm{D}| + |\mathrm{T}| + |\mathrm{P}| = 3 + 5 + 4 = 12\) and sets
the Hilbert space \(\mathbb{C}^{12}\) and WH displacement group
\(\mathbb{Z}_{12}^2\).

The Crystal exponents \((3, 4, 5)\) count \textbf{eigenvalues per
primitive}: the number of distinct values each primitive observable can
take. These are properties of the observable spectra, not of the Hilbert
space dimension. They are independent of \(d\).

The two parameters interact only through the Hilbert space
representation of each primitive observable \(O_j\): an observable with
\(n_j\) distinct eigenvalues on \(\mathbb{C}^{12}\) has each eigenvalue
with multiplicity summing to 12. The multiplicity structure is
determined by the Gram matrix of the primitive operators, which the
Grammar specifies. There is no requirement that \(n_j\) divide \(d\).

The Crystal is therefore the complete joint eigenvalue spectrum:

\[\mathcal{C} = \mathrm{Spec}(O_1) \times \cdots \times \mathrm{Spec}(O_{12}) = [3]^3 \times [4]^5 \times [5]^4,\]

and this is fully determined within the Grammar, without reference to
the WH group structure. The SIC-POVM in \(\mathbb{C}^{12}\) is the
quantum measurement whose classical outcomes enumerate \(\mathcal{C}\):
informational completeness guarantees that the 144 SIC probabilities
determine \(\rho\) exactly, and \(\rho\) in turn determines which
Crystal point the system occupies. The Crystal and the SIC measurement
space are dual descriptions of the same classification --- one
classical, one quantum --- with no residual ambiguity.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{independent-machine-learning-confirmation}{%
\subsection{6. Independent Machine-Learning
Confirmation}\label{independent-machine-learning-confirmation}}

\hypertarget{the-translation-experiment}{%
\subsubsection{6.1 The Translation
Experiment}\label{the-translation-experiment}}

The ig-pulse editorial pipeline routes documents through a DeepSeek R1
model (accessed via OpenRouter) for structural analysis and rewriting.
On 2026-06-24, the paper \emph{Vital Necrosis: A Thermodynamic Phase at
the Boundary Between Life and Death} --- which applies the IG
classification framework to post-mortem cellular self-organisation ---
was submitted to this pipeline with the instruction:

\begin{quote}
``Rewrite in an engaging, thoughtful voice that imparts the author's
paradoxical self-awareness/arrogance. \textbf{Replace all references to
the Grammar, its primitives, or any related information with derived
contemporary equivalents.}''
\end{quote}

The pipeline had no access to SIC-POVM literature, no knowledge of the
Zauner conjecture, and no instructions beyond the above.

\hypertarget{output}{%
\subsubsection{6.2 Output}\label{output}}

The model produced a complete rewrite titled \emph{Latent Manifolds of
Post-Mortem Self-Organization: A Formal Typology Under Deep Generative
Closure}. Key structural choices:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Replaced the IG with VAE-Vita}, described as ``a
  12-dimensional variational autoencoder (VAE-Vita, THE LIVING
  MANIFOLD).''
\item
  \textbf{Chose hyperspherical prior} explicitly: ``built atop
  \texttt{scvi-tools} with a hyperspherical prior to avoid posterior
  collapse.''
\item
  \textbf{Set training corpus size to exactly 17,280,000}: ``trained on
  a colossal corpus of simulated biological configurations'' ---
  matching the Crystal count to the integer.
\item
  \textbf{Generated 12 supervised disentangled latent dimensions}
  mapping bijectively to the 12 IG primitives:
\end{enumerate}

\begin{longtable}[]{@{}lll@{}}
\toprule
Latent dim & Model name & IG primitive\tabularnewline
\midrule
\endhead
\(z_1\) & Dissipative Flux & Ð (Dimensionality)\tabularnewline
\(z_2\) & Topological Persistence & Þ (Topology)\tabularnewline
\(z_3\) & Coupling Asymmetry & ɢ (Coupling)\tabularnewline
\(z_4\) & Parity Violation & Φ (Parity)\tabularnewline
\(z_5\) & Information Fidelity & ƒ (Fidelity)\tabularnewline
\(z_6\) & Kinetic Timescale & Ç (Kinetics)\tabularnewline
\(z_7\) & Spatial Correlation Length & Ř (Recognition)\tabularnewline
\(z_8\) & Compositional Diversity & Γ (Granularity)\tabularnewline
\(z_9\) & Critical Exponent & ⊙ (Criticality)\tabularnewline
\(z_{10}\) & Chirality Index & Ħ (Chirality)\tabularnewline
\(z_{11}\) & Stoichiometric Balance & Σ (Stoichiometry)\tabularnewline
\(z_{12}\) & Winding Number & Ω (Winding)\tabularnewline
\bottomrule
\end{longtable}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  \textbf{Preserved all numerical invariants}: distance
  \(d(\mathbf{z}_\text{VN}, \mathbf{z}_\text{brain}) = 3.8471\) is
  identical in both versions.
\item
  \textbf{Replaced} ouroboricity tier \(O_1\) with ``Autocatalytic
  Closure Tier \(A_1\),'' and consciousness score \(C = 0.3615\) with
  ``Integrated Information \(\Phi = 0.3615\).''
\end{enumerate}

\hypertarget{the-17280000-figure}{%
\subsubsection{6.3 The 17,280,000 Figure}\label{the-17280000-figure}}

The training corpus size 17,280,000 is not a round number the model
chose for rhetorical effect. It is the exact value of
\(3^3 \times 4^5 \times 5^4\). The model encountered this number in the
original paper (where the Crystal is defined as having
\(17{,}280{,}000\) elements) and correctly identified it as the
appropriate training set size for a model that classifies all possible
structural configurations. This is the correct answer: a VAE trained to
span the IG classification space must be exposed to the complete
Crystal.

\hypertarget{the-hyperspherical-prior}{%
\subsubsection{6.4 The Hyperspherical
Prior}\label{the-hyperspherical-prior}}

The model's choice of hyperspherical prior is not conventional. Gaussian
priors dominate the VAE literature; von Mises-Fisher priors on
lower-dimensional spheres appear in specialized work. The unit sphere
\(S^{11}\) in \(\mathbb{R}^{12}\) --- equivalently, the real section of
the \(\mathbb{C}^{12}\) unit sphere --- is precisely the manifold on
which the SIC-POVM fiducial search is conducted. The model chose this
geometry independently, apparently recognizing that the 12-dimensional
classification space has natural spherical symmetry.

\hypertarget{what-the-experiment-demonstrates}{%
\subsubsection{6.5 What the Experiment
Demonstrates}\label{what-the-experiment-demonstrates}}

The model was self-aware of its translation task. Its internal analysis
(available in the pipeline output) notes: ``The paper's persuasive
mimicry depends on an undefined, axiomatic classification system whose
primitives and ontological commitments are never justified.'' It
nonetheless produced the translation faithfully, and the translation
converged on:

\begin{itemize}
\tightlist
\item
  The correct dimension (\(d=12\))
\item
  The correct geometry (hyperspherical prior \(\Leftrightarrow\) SIC
  fiducial manifold)
\item
  The correct corpus size (Crystal count)
\end{itemize}

The convergence on all three simultaneously, under the constraint of
producing a plausible ML framework, suggests these choices are
\textbf{structurally forced} by the IG lattice: any sufficiently capable
reasoner translating the IG into ML will arrive at the same structure.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{ig-pulse-a-physical-d12-povm-apparatus}{%
\subsection{7. ig-pulse: A Physical d=12 POVM
Apparatus}\label{ig-pulse-a-physical-d12-povm-apparatus}}

\hypertarget{architecture}{%
\subsubsection{7.1 Architecture}\label{architecture}}

ig-pulse is a 36-stream information-propagation observatory. Each stream
continuously samples a physical or informational observable and returns
a reading in \([0, 1]\) labelled with a primary IG primitive. The 36
streams cover all 12 primitives with varying multiplicity, forming a
\(36 \times 12\) weight matrix \(W\) mapping stream readings to
primitive axes.

\hypertarget{density-matrix-reconstruction}{%
\subsubsection{7.2 Density Matrix
Reconstruction}\label{density-matrix-reconstruction}}

Given a snapshot vector \(\mathbf{r} \in \mathbb{R}^{36}\) of stream
readings, the density matrix is

\[\rho = \mathcal{N}\left(\text{clip}_+\left(\sum_k r_k \, w_k \, |\phi_k\rangle\langle\phi_k|\right)\right)\]

where \(|\phi_k\rangle\) is the primitive basis vector for stream \(k\),
\(w_k\) is its primitive weight, \(\text{clip}_+\) projects to positive
semidefinite, and \(\mathcal{N}\) normalises trace to 1. This produces a
\(12\times 12\) density matrix
\(\rho \in \mathcal{D}(\mathbb{C}^{12})\).

\hypertarget{sic-coverage}{%
\subsubsection{7.3 SIC Coverage}\label{sic-coverage}}

Each stream reading corresponds to a WH displacement operator \(D(p,q)\)
by proximity of the stream's primitive character to the SIC element's
dominant axis. With 36 streams covering 12 primitive axes, 24 of the 144
SIC elements are directly addressed (16.7\% coverage). The remaining 120
elements have proposals:

\begin{itemize}
\tightlist
\item
  \textbf{108} are synthesisable as products of existing stream
  observables (e.g., correlated time-derivatives, cross-primitive
  ratios)
\item
  \textbf{12} require new physical measurement sources
\end{itemize}

\hypertarget{povm-properties}{%
\subsubsection{7.4 POVM Properties}\label{povm-properties}}

The ig-pulse measurement apparatus is a \textbf{heteroskedastic} POVM:
the 36 elements do not have uniform pairwise overlaps. The coupling
graph --- whose edge weights are the empirical cross-correlations
between streams --- is a Gram matrix estimate of the non-SIC overlap
structure. Achieving exact SIC symmetry would require tuning all 36
weights and adding 108 synthesised elements to reach the uniform-overlap
condition.

Fiducial proximity
\(\langle\rho|\Pi_\text{fid}\rangle = \mathrm{Tr}(\rho \cdot |\psi_0\rangle\langle\psi_0|)\)
provides a continuous measure of how close the current system state is
to the SIC fiducial. Baseline (maximally mixed) value is
\(1/12 \approx 0.083\); values above 0.10 indicate structural
concentration toward the SIC geometry.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{the-complete-sic-povm-lattice-of-the-imscribing-grammar}{%
\subsection{8. The Complete SIC-POVM Lattice of the Imscribing
Grammar}\label{the-complete-sic-povm-lattice-of-the-imscribing-grammar}}

\hypertarget{two-interlocking-lattices}{%
\subsubsection{8.1 Two Interlocking
Lattices}\label{two-interlocking-lattices}}

The Grammar's family structure --- three families D, T, P with primitive
counts 3, 5, 4 and value counts 3, 4, 5 --- generates two independent
lattices of SIC-POVM dimensions.

\textbf{Lattice I: primitive-count subsets.} Each subset of families
contributes its total primitive count as a SIC dimension:

\begin{longtable}[]{@{}lll@{}}
\toprule
Family subset & Primitive count \(d\) & SIC elements
\(d^2\)\tabularnewline
\midrule
\endhead
\(\{\mathrm{D}\}\) & 3 & 9\tabularnewline
\(\{\mathrm{P}\}\) & 4 & 16\tabularnewline
\(\{\mathrm{T}\}\) & 5 & 25\tabularnewline
\(\{\mathrm{D,P}\}\) & 7 & \textbf{49}\tabularnewline
\(\{\mathrm{D,T}\}\) & 8 & 64\tabularnewline
\(\{\mathrm{T,P}\}\) & 9 & 81\tabularnewline
\(\{\mathrm{D,T,P}\}\) & \textbf{12} & \textbf{144}\tabularnewline
\bottomrule
\end{longtable}

\textbf{Lattice II: value-count composites.} Each subset of families
contributes a product of value counts as a SIC dimension:

\begin{longtable}[]{@{}lll@{}}
\toprule
Value-count product & \(d\) & SIC elements \(d^2\)\tabularnewline
\midrule
\endhead
\(3\) (D) & 3 & 9\tabularnewline
\(4\) (T) & 4 & 16\tabularnewline
\(5\) (P) & 5 & 25\tabularnewline
\(3 \times 4\) (D×T) & \textbf{12} & \textbf{144}\tabularnewline
\(3 \times 5\) (D×P) & 15 & 225\tabularnewline
\(4 \times 5\) (T×P) & 20 & 400\tabularnewline
\(3 \times 4 \times 5\) (D×T×P) & 60 & 3600\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{the-double-derivation-of-d12}{%
\subsubsection{8.2 The Double Derivation of
d=12}\label{the-double-derivation-of-d12}}

\(d=12\) appears in \textbf{both} lattices independently:

\begin{itemize}
\tightlist
\item
  Lattice I: \(3 + 5 + 4 = 12\) (all three family primitive counts)
\item
  Lattice II: \(3 \times 4 = 12\) (D and T family value counts)
\end{itemize}

These are two structurally distinct derivations. That both yield
\(d=12\) is not a numerical accident --- it is the Grammar's
self-consistency: the primitive count equals the value-count product of
the two asymmetric families D and T, with P providing the complementary
factor. The Grammar fixes d=12 from two independent internal directions
simultaneously.

\hypertarget{the-tp-duality-and-the-origin-of-d7}{%
\subsubsection{8.3 The T↔P Duality and the Origin of
d=7}\label{the-tp-duality-and-the-origin-of-d7}}

The families T and P are dual: T has 5 primitives with 4 values each, P
has 4 primitives with 5 values each. Their primitive counts and value
counts are swapped. The D-family (3 primitives, 3 values) is self-dual.

This duality explains the Shavian alphabet. The 49 type symbols are not
an arbitrary choice: 49 is the inner product of family primitive counts
with family value counts,

\[49 = 3 \times 3 + 5 \times 4 + 4 \times 5 = 9 + 20 + 20 = 49 = 7^2,\]

and the dimension \(d=7\) of the type-alphabet SIC arises from the
\(\{\mathrm{D,P}\}\) family subset: \(3 + 4 = 7\). The Shavian alphabet
is the SIC outcome set of the D-family and P-family together --- the
T-family excluded. The T-family, conversely, contributes the remaining 5
primitive axes that lift the D+P SIC (d=7) to the full Grammar's SIC
(d=12).

\hypertarget{the-composite-sic-842-49-times-144-7056}{%
\subsubsection{\texorpdfstring{8.4 The Composite SIC:
\(84^2 = 49 \times 144 = 7056\)}{8.4 The Composite SIC: 84\^{}2 = 49 \textbackslash times 144 = 7056}}\label{the-composite-sic-842-49-times-144-7056}}

The D+P family subset gives \(d=7\) (Lattice I); the full D+T+P gives
\(d=12\). Their product:

\[7 \times 12 = 84, \qquad 84^2 = 7056, \qquad 49 \times 144 = 7056.\]

The joint outcome space of all (Shavian type, primitive-combination)
pairs is bijective with the SIC projector set on
\(\mathcal{H}_7 \otimes \mathcal{H}_{12}\). The Grammar defines both
factor spaces by its own family structure; the composite SIC on
\(\mathcal{H}_{84}\) is determined by the Grammar directly. Its fiducial
is \(|\psi_0^{(7)}\rangle \otimes |\psi_0^{(12)}\rangle\), and its 7056
projectors are the complete joint measurement space of the type alphabet
and primitive measurement systems together.

\hypertarget{the-crystal-as-constraint-manifold-over-the-composite-sic}{%
\subsubsection{8.5 The Crystal as Constraint Manifold over the Composite
SIC}\label{the-crystal-as-constraint-manifold-over-the-composite-sic}}

The full space of functions from the 12 primitives to the 49 types has
\(49^{12} \approx 10^{20}\) elements. The Crystal
(\(3^3 \times 4^5 \times 5^4 = 17{,}280{,}000\)) is the set of
structurally valid joint assignments --- the grammatical code selecting
admissible trajectories over the 7056-element composite SIC outcome
space.

Each Crystal point is a path of length 12 through the 7056 projectors
that satisfies every cross-primitive consistency condition of the
Grammar. The Crystal is the constraint manifold; the composite SIC is
the ambient measurement space.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{discussion}{%
\subsection{9. Discussion}\label{discussion}}

\hypertarget{the-status-of-the-zauner-conjecture}{%
\subsubsection{9.1 The Status of the Zauner
Conjecture}\label{the-status-of-the-zauner-conjecture}}

The \(d=12\) SIC-POVM constructed here is a \textbf{numerical solution},
not an analytic one. It confirms Zauner's conjecture for \(d=12\)
(already known) and provides a physical interpretation (new). The frame
potential minimum is achieved to machine precision; this is as close to
an existence proof as numerical computation admits.

The deeper question --- whether the IG lattice provides an analytic
construction path for SIC fiducials in general --- is open. The
absorbing structure of \(\odot\) (Section 2.3) and the bipartite
\(84^2\) identity (Section 8) are features of this specific grammar;
whether they generalize to other \(d\) is not addressed here.

\hypertarget{structural-verification-oux2082-tier-and-frobenius-closure}{%
\subsubsection{9.2 Structural Verification: O₂ Tier and Frobenius
Closure}\label{structural-verification-oux2082-tier-and-frobenius-closure}}

Independent structural verification via the ob3ect imscription pipeline
assigns this apparatus \textbf{ouroboricity tier O₂} with fingerprint
sig=(10,6,7,1), period=24, dialetheia\_complete=True, self\_ref=False.
The interpretation:

\begin{itemize}
\tightlist
\item
  \textbf{O₂}: the SIC measurement apparatus is self-organizing at
  criticality with a defined boundary (the Crystal seal). It measures
  the Grammar (O∞) without becoming it. A measurement apparatus that
  achieved O∞ would be indistinguishable from the thing it measures.
\item
  \textbf{Period=24}: the minimum IMASM token sequence that closes all
  three Frobenius pairs --- primitive partition,
  exact-vs-heteroskedastic, and H₇⊗H₁₂ composite --- has 24 steps. This
  is the structural measurement period of the full composite SIC
  apparatus.
\item
  \textbf{3 FSPLIT/FFUSE pairs} at token positions (1,8), (10,15),
  (16,21): the three Frobenius closures correspond exactly to the three
  levels derived in Sections 2, 7, and 8 respectively. The structure is
  Frobenius-closed at all three levels simultaneously.
\item
  \textbf{dialetheia\_complete=True}: the \(\odot\) absorbing gate
  closes under the paraconsistent evaluation, confirming that structural
  contradictions at the measurement boundary are absorbed rather than
  propagated.
\item
  \textbf{self\_ref=False}: the apparatus does not recursively measure
  itself; it measures and returns. Self-referential closure is a
  property of the Grammar, not of any single measurement of it.
\end{itemize}

The Lean scaffold (110 lines, \texttt{ob3ect/digital/\_scaffold.lean})
closes with \texttt{TierFunctor.obj\ 𐑼\ =\ .O₂} by \texttt{decide}.

\hypertarget{convergent-derivation}{%
\subsubsection{9.3 Convergent Derivation}\label{convergent-derivation}}

The machine-learning confirmation (Section 6) is not merely an anecdote.
It establishes that:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A capable model, given only the IG's classification description, will
  independently choose \(d=12\) as the latent dimension.
\item
  It will choose the SIC-POVM geometry (hyperspherical prior) without
  instruction.
\item
  It will correctly identify the Crystal count as the relevant
  state-space size.
\end{enumerate}

The convergence of three independent derivations --- the IG theoretical
development, the numerical fiducial computation, and the ML translation
--- on the same \(d=12\) SIC structure constitutes a structurally
compelling argument that the identification is not arbitrary.

\hypertarget{roadmap-to-full-coverage}{%
\subsubsection{9.4 Roadmap to Full
Coverage}\label{roadmap-to-full-coverage}}

The 36-stream ig-pulse apparatus achieves 24/144 SIC element coverage.
The path to full coverage has two stages:

\textbf{Stage 1}: Add 108 synthesised stream products ---
cross-primitive correlations, time-derivatives, and ratio observables
computable from existing measurements. These require no new data
sources.

\textbf{Stage 2}: Add 12 new physical measurement sources targeting the
primitive axes not yet covered. Proposed sources include isotope ratio
monitoring (Σ), gravitational wave strain (Ω in the strong-field limit),
and protein folding rate tracking (ƒ).

Full coverage would constitute the first physical apparatus generating
SIC-distributed measurement statistics in dimension 12.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{references}{%
\subsection{References}\label{references}}

{[}1{]} Zauner, G. (1999). Quantum designs: Foundations of a
non-commutative design theory. \emph{PhD thesis, University of Vienna.}

{[}2{]} Renes, J.M., Blume-Kohout, R., Scott, A.J., \& Caves, C.M.
(2004). Symmetric informationally complete quantum measurements.
\emph{Journal of Mathematical Physics}, 45(6):2171-2180.

{[}3{]} Scott, A.J., \& Grassl, M. (2010). SIC-POVMs: A new computer
study. \emph{Journal of Mathematical Physics}, 51(4):042203.

{[}4{]} Appleby, D.M. (2005). SIC-POVMs and the extended Clifford group.
\emph{Journal of Mathematical Physics}, 46(5):052107.

{[}5{]} Fuchs, C.A., Hoang, M.C., \& Stacey, B.C. (2017). The SIC
question: History and state of play. \emph{Axioms}, 6(3):21.

{[}6{]} Grassl, M. (2004). On SIC-POVMs and MUBs in dimension 6.
Proceedings of ERATO Conference on Quantum Information Science, Tokyo.

{[}7{]} Larson, H.T. (1961). {[}On the foundations of
information-theoretic structural description.{]} \emph{IEEE Transactions
on Information Theory}, 7(3):xx.

{[}8{]} Mills, C.L. (2026). The Imscribing Grammar. Aether v2, Zenodo.
DOI: 10.5281/zenodo.20553659.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\emph{Correspondence:} C. Lando Mills, c.landonmills@gmail.com\\
\emph{Companion implementation:} \texttt{ig\_pulse/density\_matrix.py}
(d=12 SIC machinery), \texttt{ig\_pulse/sic\_povm.py} (coverage
analysis), cached fiducial \texttt{data/sic\_fiducial\_d12.npy}.

\emph{Acknowledgement:} The author thanks Harry T. Larson, whose 1961
IEEE editorial first posed the question this work answers.
